---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r chunkopts, eval = TRUE, echo = FALSE}
knitr::opts_chunk$set(eval = TRUE, echo = TRUE, message = FALSE, warning = TRUE, tidy = FALSE, cache = TRUE, include = TRUE, dpi = 72, fig.width = 9, fig.height = 9, fig.align = "center", results = "markup",
  collapse = TRUE,
  comment = "#>",
  fig.path = "vignettes/README-"
)

```

# SOmap, SO-CPR data, and other things

Ben Raymond, AAD

November 2019

See also the SOmap section of the notes from [this R workshop](https://scar.github.io/EGABIcourse19/mapping.html#somap).


## Preparation

Install some packages if necessary:

```{r}
have_pkg <- function(p) requireNamespace(p, quietly = TRUE)
github_pkgs <- c("AustralianAntarcticDivision/blueant", "AustralianAntarcticDivision/SOmap")
needed <- sapply(github_pkgs, function(z) !have_pkg(basename(z)))
if (any(needed)) {
  if (!have_pkg("remotes")) install.packages("remotes")
  for (p in github_pkgs[needed]) remotes::install_github(p)
}

pkgs <- c("data.table", "dplyr", "ggplot2", "lubridate")
needed <- !sapply(pkgs, have_pkg)
if (any(needed)) install.packages(pkgs[needed])

```

## Download the SO-CPR data and read it in

```{r}
library(blueant)
library(data.table)
library(lubridate)

data_dir <- tempdir()
status <- bb_get(sources("SO-CPR"), local_file_root = data_dir)

x <- fread(status$files[[1]]$file, data.table = FALSE)
x$date <- dmy(x$Date)

## the data looks like
dplyr::glimpse(x)

```

## Semi-automatic maps

Just a base map:

```{r}
library(SOmap)

SOmap(trim = -40)

```

To which we can add stuff:

```{r}
SOmap(trim = -40)
SOplot(x$Longitude, x$Latitude)

```

## A subset of the data

An East Antarctic subset of the data:

```{r}
xsub <- x[x$Longitude >= 50 & x$Longitude <= 150, ]

SOmap(trim = -40)
SOplot(xsub$Longitude, xsub$Latitude)
```

Maybe better on something other than full polar-stereo projection?

```{r}
SOmap_auto(xsub$Longitude, xsub$Latitude, input_lines = FALSE)

```

## More complicated

```{r}
library(ggplot2)
## make a base ggplot2 map
p <- SOgg(SOmap(trim = -40))

## get projected coordinates for our data
px <- SOproj(x$Longitude, x$Latitude, target = p$projection)
x <- cbind(x, coordinates(px))

## plot the map and add points with colour by ship code
plot(p) + geom_point(data = x, aes(colour = `Ship Code`))

```

## Other R stuff

The [SOmap package](https://github.com/AustralianAntarcticDivision/) is part of a larger suite of packages to support science in the Southern Ocean and Antarctica. See the [SCAR/rOpenSci initiative](https://scar.github.io/ropensci/) for more information.

The [RAADtools](https://github.com/AustralianAntarcticDivision/raadtools) package provides tools for working with a suite of (mostly Southern Ocean) environmental data.

Using the last 1000 rows only as an example, we use `raadtools` to extract time- and location-matched sea surface height anomaly data:

```{r}
library(raadtools)

xsub <- tail(x, 1000)
xsub$SSHA <- extract(readssh, tail(x[, c("Longitude", "Latitude", "date")], 1000), anomaly = TRUE)
```

Plotted against total abundance:

```{r}
ggplot(xsub, aes(SSHA, `Total abundance`)) + geom_point() + theme_bw()

```

